---
title: "Week2_reproducible_research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading and preprocessing the data
Show any code that is needed to

    Load the data (i.e. read.csv())

    Process/transform the data (if necessary) into a format suitable for your analysis
    
```{r}
library(readxl)
library(data.table)
library(ggplot2)
library(dplyr)
#read in csv 
activity <-fread("./activity.csv")
```

#What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
```{r}
#calculate mean steps per day by aggregating
totalstepsday <-  aggregate(steps ~ date, activity, FUN = sum)

#make a histogram
ggplot(totalstepsday, aes(x=steps)) +
geom_histogram()



```

```{r}

#calculate mean steps per day by summarize
totalstepsday %>% summarize(mean(steps))

```

```{r}

#calculate median steps per day by summarize
totalstepsday %>% summarize(median(steps))
```

#What is the average daily activity pattern?

```{r}
library(lubridate)

#Make a time series plot (i.e. type = "l"\color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
meansteps<- activity %>%
  group_by(interval) %>%
  summarize(meansteps = mean(steps, na.rm = TRUE))


plot(meansteps$meansteps ~ meansteps$interval, type = "l", xlab = "5 min interval", ylab = "average steps per day"
    )
```

```{r}
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
meansteps %>% subset(meansteps == max(meansteps))
```

#Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA\color{red}{\verb|NA|}NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA\color{red}{\verb|NA|}NAs)
summary(is.na(activity$steps))
#2304
```
```{r}
#Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

#use mean of the day divided by the total of intervals
activity <- activity %>%
  ungroup()%>%
  mutate(steps_day = as.numeric(mean(steps, na.rm=TRUE))) %>%
    group_by(date) %>%
         mutate(interval_n = as.numeric(n())) %>% 
mutate(steps_imp = dplyr::case_when(
  is.na(steps)== FALSE ~ as.numeric(steps),
  is.na(steps)== TRUE ~ steps_day/interval_n))

head(activity)
```
```{r}
#Create a new dataset that is equal to the original dataset but with the missing data filled in.
activity_2 <- activity %>%
  select(steps_imp, date, interval) %>%
  rename(steps = steps_imp)

summary(is.na(activity_2$steps))
```

```{r}
#Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

#calculate mean steps per day by aggregating
totalstepsday_2 <-  aggregate(steps ~ date, activity_2, FUN = sum)

#make a histogram
ggplot(totalstepsday, aes(x=steps)) +
geom_histogram()
```

```{r}

#calculate mean steps per day by summarize
totalstepsday_2 %>% summarize(mean(steps))

#mean(steps)
#10766.19	
#before it was much higher

```

```{r}

#calculate median steps per day by summarize
totalstepsday_2 %>% summarize(median(steps))

#median(steps)
#10765	
#not that different, the median is not too much influenced

```

#Are there differences in activity patterns between weekdays and weekends?

```{r}
#Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
activity_2 <- activity_2 %>%
  mutate(weekday= weekdays(date)) %>%
  mutate(week = case_when(weekday != "zaterdag"&
                            weekday != "zondag"~ "weekday",
                          weekday == "zaterdag" | 
                            weekday == "zondag" ~ "weekend"))
table(activity_2$week)
table(activity_2$weekday)
```

```{r}
#Make a panel plot containing a time series plot (i.e. type = "l"\color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

#average steps per interval and make two plots for weekend and weekdays
meansteps_2_week<- activity_2 %>%
  subset(week == "weekday")%>%
  group_by(interval) %>%
  mutate(meansteps = mean(steps, na.rm = TRUE))

meansteps_2_weekend<- activity_2 %>%
  subset(week == "weekend")%>%
  group_by(interval) %>%
  mutate(meansteps = mean(steps, na.rm = TRUE))

par(mfrow=c(2,1))

week <-plot(meansteps_2_week$meansteps ~ meansteps_2_week$interval, type = "l", xlab = "5 min interval", ylab = "average steps per interval weekend"
    )

weekend <- plot(meansteps_2_weekend$meansteps ~ meansteps_2_weekend$interval, type = "l", xlab = "5 min interval", ylab = "average steps per interval weekend"
    )


```


